Paradigms in Physics 2.0

### 2017-06-27 paper.tex
input from Corinne and Ethan

@@

\subsubsection{Challenges in Contemporary Physics}
Climate change and renewable energy are major contemporary
challenges. Students are inspired to engage with these challenges, and
physics will play a critical role in addressing these challenges.
Former Energy Secretary Steve Chu recently noted that ``physics majors
seem to be increasing,'' and pointed to millennials' interests in the
risks of climate change, clean energy, and
sustainability~\cite{kramer2016gathering}.  However, most majors get
little or no interaction with such challenges in their physics
classes.

We have introduced a course that fits seamlessly into the physics
major sequence, that focuses on the contemporary challenges of
sustainable energy and climate change. The course will be a bridge
between introductory physics and upper-division classes for both
physics majors and students from related disciplines such as chemistry
and engineering. The challenges theme will motivate the introduction
of thermal physics and quantum mechanics as well as the development of
physical reasoning skills. We anticipate that an emphasis on the role
of physics in overcoming societal challenges will help us recruit and
retain a greater number of physics majors, especially women and
under-represented minorities.

At Oregon State, as at many other universities, a course in Modern
Physics traditionally bridges the lower and upper divisions and serves
as an introduction to quantum mechanics. Such courses typically follow
a historical narrative, celebrating the achievements of men like
Planck, Einstein, Rutherford, Bohr, Bragg, Millikan, de Broglie,
Pauli, Schroedinger, and Dirac. Indeed, it is a wonderful and exciting
story - and inspiring for those who imagine themselves as
reincarnations of these men. However, we suggest that this historical
narrative does not attract a wide cross-section of participants to
physics. Instead, we have introduced a forward-looking narrative that
emphasizes the relevance of physics today unconstrained by the details
of the great transition from 1905 to 1945.

Each physics topic is linked to contemporary challenges. An
understanding of electromagnetic waves is required to understand how
light interacts with vibrational modes of CO$_2$. The equipartition
theorem explains the heat energy stored in the ocean. Quantum
mechanics explains the blackbody spectrum, semiconductor electronic
structure, and solar photovoltaics. The Carnot cycle convinces us that
a 95\%-efficient gas furnace cannot compete with an efficient heat
pump.

The new perspectives associated with this course, along with an
emphasis on expanding students' reasoning skills, will strengthen the
connection between lower- and upper-division physics classes. Students
will learn the importance of approximate calculations (Fermi
estimates) that establish reasonable bounds on the exact
answer. Students will learn that equations help us interpret the
physical world, i.e. equations are used for much more than mindless
algebraic manipulations.  Students will learn how integration and
differentiation can be applied to real data - not just the perfect
functions that are taught in math class. Students will learn how
graphing of relationships is key tool for physicists. The utility of
these skills will be firmly established by application to contemporary
challenges.

@@

\section{Outcomes}
\emph{How has it gone?}

\subsection{Faculty vote}
\emph{Unanimous result.}
\subsection{Transition year}
\emph{Gradual transition, getting courses taught.}

\section{Future research}
\emph{Document learning trajectories, new courses.}

>>>>

\emph{Faculty vote}

@@

\subsubsection{Challenges in Contemporary Physics}

In ``The Physics of Contemporary Challenges'' we prepare students to
apply physics concepts and physical reasoning skills to sustainable
energy issues, climate change mechanisms, space exploration and
puzzles in fundamental physics. These ``real-world'' topics are chosen
for either the societal need (energy and climate), and/or the human
need to explore (space and fundamental physics). By prioritizing
inclusion of the most engaging challenges, we aim to attract and
retain as many potential physics majors as possible.

While contemporary challenges determine the narrative flow of the
course, physics concepts and physical reasoning skills are the main
substance. Students are introduced to thermodynamics, statistical
mechanics, electromagnetic radiation, quantum mechanics, and modern
experimental physics---in each case motivated by one or more
``contemporary challenges''. Together with new physics concepts,
students are given new reasoning tools, such as the equipartition
theorem, the quantum-classical correspondence principle, order of
magnitude calculations, simplifying assumptions, and numerical
integration. Mathematically rigorous derivations are only briefly
discussed in class, with additional details offered in online video
format. In summary, we emphasis the physics concepts and reasoning
skills that allow professional physicists to quickly/quantitatively
make an initial assessment of a complex problem.

The students engage with the material by solving daily quiz questions
in class, solving weekly homework problems, performing a modern
physics experiment, and writing a term paper. The daily quiz questions
help clarify the goal of each lecture. The weekly homework questions
are directly related to contemporary challenges are require a variety
of estimation techniques. The modern physics experiment, performed
with a low cost kit (LEDs, diffraction grating, rulers, volt meter
etc.) emphasizes a quantitative approach to observation and explores
issues of experimental uncertainty. Lastly, the term paper allows
students to dive deeper into a particular contemporary challenge. The
term paper tests student's ability to make their own initial
assessment of a complex problem, and communicate their ideas with
diagrams, graphs and estimations.

%% Climate change and renewable energy are major contemporary
%% challenges. Students are inspired to engage with these challenges, and
%% physics will play a critical role in addressing these challenges.
%% Former Energy Secretary Steve Chu recently noted that ``physics majors
%% seem to be increasing,'' and pointed to millennials' interests in the
%% risks of climate change, clean energy, and
%% sustainability~\cite{kramer2016gathering}.  However, most majors get
%% little or no interaction with such challenges in their physics
%% classes.

%% We have introduced a course that fits seamlessly into the physics
%% major sequence, that focuses on the contemporary challenges of
%% sustainable energy and climate change. The course will be a bridge
%% between introductory physics and upper-division classes for both
%% physics majors and students from related disciplines such as chemistry
%% and engineering. The challenges theme will motivate the introduction
%% of thermal physics and quantum mechanics as well as the development of
%% physical reasoning skills. We anticipate that an emphasis on the role
%% of physics in overcoming societal challenges will help us recruit and
%% retain a greater number of physics majors, especially women and
%% under-represented minorities.

%% At Oregon State, as at many other universities, a course in Modern
%% Physics traditionally bridges the lower and upper divisions and serves
%% as an introduction to quantum mechanics. Such courses typically follow
%% a historical narrative, celebrating the achievements of men like
%% Planck, Einstein, Rutherford, Bohr, Bragg, Millikan, de Broglie,
%% Pauli, Schroedinger, and Dirac. Indeed, it is a wonderful and exciting
%% story - and inspiring for those who imagine themselves as
%% reincarnations of these men. However, we suggest that this historical
%% narrative does not attract a wide cross-section of participants to
%% physics. Instead, we have introduced a forward-looking narrative that
%% emphasizes the relevance of physics today unconstrained by the details
%% of the great transition from 1905 to 1945.

%% Each physics topic is linked to contemporary challenges. An
%% understanding of electromagnetic waves is required to understand how
%% light interacts with vibrational modes of CO$_2$. The equipartition
%% theorem explains the heat energy stored in the ocean. Quantum
%% mechanics explains the blackbody spectrum, semiconductor electronic
%% structure, and solar photovoltaics. The Carnot cycle convinces us that
%% a 95\%-efficient gas furnace cannot compete with an efficient heat
%% pump.

%% The new perspectives associated with this course, along with an
%% emphasis on expanding students' reasoning skills, will strengthen the
%% connection between lower- and upper-division physics classes. Students
%% will learn the importance of approximate calculations (Fermi
%% estimates) that establish reasonable bounds on the exact
%% answer. Students will learn that equations help us interpret the
%% physical world, i.e. equations are used for much more than mindless
%% algebraic manipulations.  Students will learn how integration and
%% differentiation can be applied to real data - not just the perfect
%% functions that are taught in math class. Students will learn how
%% graphing of relationships is key tool for physicists. The utility of
%% these skills will be firmly established by application to contemporary
%% challenges.

@@

\section{Summary}
\emph{Rearticulate big ideas. Shadowing}

\section{Future research}
\emph{Document learning trajectories, new courses.  pre-/post-,
  understanding sense-making.}
